### 2018-12-23 papers/journal/experiments.tex
Added encodings for recahabilty for BMC implementation
added QBF section and attached Synthesis encodings to it
restructure introduction and started experiments section. Added references to the bio part of the intro section.

We have developed three different tools (satVTS, smtVTS, synthVTS) for implementation of proposed encodings. All the experiments were done on a machine with Intel(R) Core(TM) i3-4030U CPU @ 1.90GHz processor and 4GB RAM with 30 min (1800 sec) timeout.

\input{result-table}

\subsection{SATVTS (Bounded model checker based tool)} 
The first tool is implemented using model checking tool for C program CBMC~\cite{clarke2004tool}.
%
CBMC is a Bounded Model Checker for C and C++ programs. 
%
Unlike a general model checker~\cite{cimatti2002nusmv}, CBMC checks the verified property for all the possible states, only up to a certain “depth,” a parametric limit on the size of the model.
%
The depth of model is specified by the user and can be  incrementally increased to a desired value.

@@

\subsection{SMTVTS/pyVTS (SMT based tool)}
	We have implemented the encodings for each variants using the python interface of Z3 in a tool (MAA). 
%
Our tool allows the user to choose a model and the size
of the network besides other parameters like connectivity and number of parallel edges. 
%		
It uses Z3 Python interface to build the needed constraints and applies Z3 solver on the constraints to find a model ( a satisfying assignment that respects the constraints). 
%
This tool also translates the satisfying model found by Z3 into
a VTS and presents a visual output to the user in form of annotated graph. 
%
We also visually report the dropped edges required to disconnect the graph, it gives information about the connectivity of the graph.
%
To illustrate usability of our tool in the last column of the table 1, we present the minimum connectedness needed for the different variants after applying our tool for sizes from 2 to 10. 
%
We found no graph for the variant A with constraints Ann and Aen. 
%
Replacing constraint on the node of Ann with Anb (variant B)
does not affect the outcome. 
%
If we allow every present molecule to stay active
(Ann) but constraint the edge by a boolean function (Aeb) the resultant VTS has to be at least 3-connected. 
%
Similarly, the results for the other cases are presented in the table.

In Table~\ref{tab:qf-grabh}, we present the running times for the sear
ch of
VTSs of sizes 2 to 10 that satisfy the variants.
and compare with our old encoding
({Old-e}) from~\cite{shukla}.
%
For the comparison between both encodings we have fixed the total
number of molecules to be $|M| = 2|N|$ for $ |N|> 2$ and
$|M| = 2|N| + 1$ for $|N| = 2$.
%
For each variant, we fix maximum number of parallel
edges to 2.
%
In the table we have shown comparison for specific connectivity, for
example variant A is checked against any graph with connectivity 2,
variant B with connectivity 3 similarly for the rest of the Variants.
%

@@

\subsection{SynthVTS (QBF based synthesis tool)} 
We have implemented the encodings in a tool
called~\ourtool\footnote{{\url{https://github.com/arey0pushpa/pyZ3}}}.
%
The tool takes a partially defined VTS as input in a custom designed
input language.
%
The input is then converted to the constraints over VTS. 
%
The tool can not only synthesize the above-discussed queries, but also their
combinations.
%
For example, our tool can modify labels of nodes or edges while
learning activity functions.
%
Our tool is developed in C++ and uses~\zthree~\cite{z3} infrastructure for
processing formulas. 
%
Since some of the formulas involve alternation of quantifiers over
Boolean variables Z3 is not a suitable choice for those examples.
%
We translate the formulas created by Z3 tool into a standard
QDIMACS~\cite{qdimacs} format and use as an input for QBF solvers. 
%
We use~\depqbf~\cite{lonsing2010depqbf} for solving of QBF formulas. 
%
Our tool includes about 7000 lines of code.

>>>>

%\externaldocument{sat-table}

%We have developed four different tools (\sattool, \smttool, \qbftool, \ourtool) for implementation of proposed encodings.
%
We have developed the {\vtstool} tool to conjecture the properties of the VTS. 
%
The toolchain is divided into three different blocks. 
%
Checking existential condition, checking the universal condition and performing synthesis on the incomplete VTS. 
%
The existential condition is subdivided into two sub-block, using bounded model checking and using constraint solving. 
%
We will refer to existential subblocks as {\sattool} and {\smttool}, universal block as {\qbftool} and synthesis block as {\ourtool}.

\subsection{Implementation}
{\sattool} is based on the architecture of a Bounded model checker (CBMC) and uses an enumeration of paths for encoding the reachability problem.
%
{\smttool} is based on the SMT solver Z3 and uses the efficient encoding of finding a least fixed point to solve reachability problem. 
%
{\qbftool} is based on the QBF encoding of the universal condition and uses {\depqbf} solver to solve the generated QBF. 
%
The {\ourtool} is a synthesis tool for synthesizing incomplete information in the input VTS.
%
All the experiments were done on a machine with  Intel(R) Core(TM) i7-2600 CPU @ 3.40GHz processor and 16GB RAM with 30 min (1800 sec) timeout.

%\input{sat-table}

%\input{tab:satqbf-grabh}
%\ExecuteMetaData[result-table.tex]{tab:satqbf-grabh}

\textbf{{\sattool} (Bounded model checker based tool)}:
The {\sattool} tool is implemented using model checking tool for C program CBMC~\cite{clarke2004tool}.
%
CBMC is a Bounded Model Checker for C and C++ programs. 
%
Unlike a general model checker~\cite{cimatti2002nusmv}, CBMC checks the verified property for all the possible states, only up to a certain “depth,” a parametric limit on the size of the model.
%
The depth of model is specified by the user and can be  incrementally increased to a desired value.

@@

\textbf{{\smttool} (SMT based tool)}:
	We have implemented the encodings for each variants using the python interface of Z3 in a tool {\smttool}. 
%
Our tool allows the user to choose a model and the size
of the network besides other parameters like connectivity and number of parallel edges. 
%		
It uses Z3 Python interface to build the needed constraints and applies Z3 solver on the constraints to find a model ( a satisfying assignment that respects the constraints). 
%
This tool also translates the satisfying model found by Z3 into
a VTS and presents a visual output to the user in form of annotated graph. 
%
We also visually report the dropped edges required to disconnect the graph, it gives information about the connectivity of the graph.
%
To illustrate usability of our tool in the last column of the table 1, we present the minimum connectedness needed for the different variants after applying our tool for sizes from 2 to 10. 
%
We found no graph for the variant A with constraints Ann and Aen. 
%
Replacing constraint on the node of Ann with Anb (variant B)
does not affect the outcome. 
%
If we allow every present molecule to stay active
(Ann) but constraint the edge by a boolean function (Aeb) the resultant VTS has to be at least 3-connected. 
%
Similarly, the results for the other cases are presented in the table.

\input{smt-table}

In Table~\ref{tab:smt-grph}, we present the running times for the sear
ch of
VTSs of sizes 2 to 10 that satisfy the variants.
and compare with our old encoding
({Old-e}) from~\cite{shukla}.
%
For the comparison between both encodings we have fixed the total
number of molecules to be $|M| = 2|N|$ for $ |N|> 2$ and
$|M| = 2|N| + 1$ for $|N| = 2$.
%
For each variant, we fix maximum number of parallel
edges to 2.
%
In the table we have shown comparison for specific connectivity, for
example variant A is checked against any graph with connectivity 2,
variant B with connectivity 3 similarly for the rest of the Variants.
%

@@

\textbf{{\qbftool} (QBF based synthesis tool):}

\input{synth-table}
We have implemented the encodings in a tool
called~\ourtool\footnote{{\url{https://github.com/arey0pushpa/pyZ3}}}.
%
The tool takes a partially defined VTS as input in a custom designed
input language.
%
The input is then converted to the constraints over VTS. 
%
The tool can not only synthesize the above-discussed queries, but also their
combinations.
%
For example, our tool can modify labels of nodes or edges while
learning activity functions.
%
Our tool is developed in C++ and uses~\zthree~\cite{z3} infrastructure for
processing formulas. 
%
Since some of the formulas involve alternation of quantifiers over
Boolean variables Z3 is not a suitable choice for those examples.
%
We translate the formulas created by Z3 tool into a standard
QDIMACS~\cite{qdimacs} format and use as an input for QBF solvers. 
%
We use~\depqbf~\cite{lonsing2010depqbf} for solving of QBF formulas. 
%
Our tool includes about 7000 lines of code.

### 2018-12-23 papers/journal/intro.tex
removed first subsection of intro
restructure introduction and started experiments section. Added references to the bio part of the intro section.

@@

\textbf{Major paths in the vesicle transport network}:
Molecules traverse the cell in a series of vesicles, and there are two major such routes of transport in all eukaryotic cells:
\begin{enumerate}
	\item the secretory route, which takes proteins from the ER, their site of production, to the plasma
	membrane, from where they are secreted out of the cell.
	\item the endocytic route, which takes proteins from the outside of the cell through the plasma
	membrane to the endocytic compartments, where they are digested.
\end{enumerate}

Other paths are used for cross talk between the secretory and the endocytic routes.
%
There are also cyclic paths for recycle of proteins which are necessarily dislodged during the process of vesicle traffic back to their original compartments.

\textbf{Biological Model of VTS}
%
A characteristic feature of eukaryotic cells is the presence of multiple internal membrane bound organelles. 
%
These organelles exchange molecules amongst themselves in small membrane bound packets called vesicles. 
%
The recognition of cargo molecules to be loaded onto these vesicles and the recognition of the correct target organelles for each vesicle is performed by an intricate network of molecules called the vesicle transport system []. 
%
We call the set of internal organelles, together with the molecular routes that connect them through vesicle exchange, the vesicle transport network. 
%
Although vesicle mediated traffic was discovered decades ago [17], the picture of the vesicle transport network is far from complete; we do not yet know how many of the cellular proteins reach their resident organelles within the cell, and new vesicle routes are being discovered every year [1, 3].

Completing the vesicle traffic network is a difficult task for many reasons: (1) the core vesicle transport network, which consists of a secretory and an endocytic route, is conserved across all eukaryotes, but the traffic network in different organisms [6, 12, 14], and even in different cell types within an organism can be different [10, 11], (2) although the basic traffic machinery is the same for all vesicle fusion events, the details of regulation can be different [15, 16], (3) behaviour of traffic molecules in vitro is different from their behaviour inside cells [9], (4) molecules can have redundant routes within the cell [7,8], and (5) it can be difficult to distinguish between the direct and indirect effects of experiments involving knock-downs or knock outs of traffic molecules [2, 13].

Nonetheless, completing the vesicle traffic network for different organisms is very useful. Many diseases are caused by the malfunction of the vesicle traffic network. Knowledge of the complete network would be helpful in identifying the root causes.[4,5]. At a more basic level, having complete pictures of the vesicle traffic system for various organisms and various cell-types would allow comparative studies, and therefore would allow the deduction of modes of evolution of the form of the traffic network, and would also allow us to decipher features such as parts of the traffic system that are unchanging, and therefore likely to be its core, and parts that are more plastic [14].

In this paper, using constraints on global vesicle traffic network topology due to local molecular interactions, we take incomplete pictures of vesicle traffic networks as inputs and output various completed versions which can then be tested experimentally.

>>>>

%\textbf{Components of the vesicle transport network}:
A characteristic feature of eukaryotic cells is the presence of multiple internal membrane-bound organelles. 
These organelles exchange molecules amongst themselves in small membrane-bound packets called vesicles. 
%
The recognition of cargo molecules to be loaded onto these vesicles and the recognition of the correct target organelles for each vesicle is performed by an intricate network of molecules called the vesicle transport system~\cite{alberts2002molecular}. 
%
We call the set of internal organelles, together with the molecular routes that connect them through vesicle exchange, the vesicle transport network. 
%
Vesicle traffic underpins nearly every aspect of eukaryotic cellular physiology, including
human cellular physiology.
%
Understanding how this system function is, therefore,
one of the central challenges of cell biology.

\vspace{0.5cm}

@@

\textbf{Major paths in the vesicle transport network}:
Molecules traverse the cell in a series of vesicles, and there are two major such routes of transport in all eukaryotic cells:
\begin{enumerate}
	\item The secretory route takes proteins from the ER, their site of production, to the plasma membrane, from where they are secreted out of the cell. Cargo proteins leave the ER in COP2 coated vesicles which fuse with the Golgi apparatus. The golgi apparatus is where proteins get modified, for example by the addition of carbohydrate side-chains. Subsequently, proteins destined for secretion are packaged into clathrin coated vesicles which fuse with the plasma membrane, thus releasing their contents to the outside of the cell\cite{alberts2002molecular}.
	\item the endocytic routetakes proteins from the outside of the cell through the plasma membrane to the endocytic compartments, where they are digested. Cargo from the outside of the cell is taken in in the form of clathrin coated vesicles which form at the plasma membrane. These vesicles fuse amongst themselves in a process called homotypic fusion to generate the early endosomal compartments. Early endosomes undergo changes in their composition due to ongoing vesicle traffic and maturee into late endosomes. This maturation is concomitant with a switch the Rab protein they are associated with; the Rab5 associated early endosomes mature into Rab7 associated late endosomes\cite{rink2005rab}. Late endosomes then fuse with lysosomes, where all its contents get digested\cite{pryor2009delivery}.
\end{enumerate}

\textbf{Existential and universal condition:}
Other paths are used for cross talk between the secretory and the endocytic routes. For example, vesicles are sent from the TGN (Trans-Golgi network) to the early and the late endosomes to transport enzymes, and vesicles are sent back from the endocytic compartments to the TGN to recycle sorting receptors\cite{progida2016bidirectional}. Also recently, evidence for unconventional vesicle-mediated secretory routes have been found, which bypass the Golgi. The molecules involved in these routes are as yet unknown\cite{nickel2018unconventional}. 
%

\begin{table}
	\begin{center}
		\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
			\hline
			Nodes & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\ \hline
			Graphs & 0 & 0 & 0 & 1 & 2 & 15 & 121 & 2159 & 68715 & 3952378 \\ 
			\hline
		\end{tabular}
		\label{tab:threec}
		\caption{Number of simple 3-edge-connected unlabelled N-node graphs.}
	\end{center}	
\end{table}

\textbf{Difficulty of the problem:}
The analysis of vesicle traffic systems is a difficult problem
because of the combinatorial scaling of possible traffic topologies and regulatory rules. 
%
For example, we might want to check some conjecture of interest for all 3-connected graphs and
all possible variations of SNARE regulation rules. 
%
The number of graphs of specified connectivity grows exponentially with the number of nodes: Table~\ref{tab:threec} shows how many 3-edge-connected graphs~\cite{a052448-oeis} exist (without parallel or self edges) as node number N increases.

\textbf{Synthesis of VTS:}
Although vesicle-mediated traffic was discovered decades ago~\cite{wells2005discovery}, the picture of the vesicle transport network is far from complete; we do not yet know how many of the cellular proteins reach their resident organelles within the cell, and new vesicle routes are being discovered every year ~\cite{nickel2018unconventional,weill2018toolbox}.

Completing the vesicle traffic network is a difficult task for many reasons: (1) the core vesicle transport network, which consists of a secretory and an endocytic route, is conserved across all eukaryotes, but the traffic network in different organisms \cite{richardson2015evolutionary,nishimoto2009differential,barlow2017seeing}, and even in different cell types within an organism can be different \cite{stoops2014trafficking,zhou2015arp2}, (2) although the basic traffic machinery is the same for all vesicle fusion events, the details of regulation can be different \cite{davletov2007regulation,di2010calcium}, (3) behaviour of traffic molecules in vitro is different from their behaviour inside cells \cite{furukawa2014multiple}, (4) molecules can have redundant routes within the cell \cite{shimizu2014compensatory,nakatsukasa2014nutrient}, and (5) it can be difficult to distinguish between the direct and indirect effects of experiments involving knock-downs or knock outs of traffic molecules \cite{hirst2004epsinr,mishev2013small}.

Nonetheless, completing the vesicle traffic network for different organisms is very useful. Many diseases are caused by the malfunction of the vesicle traffic network. Knowledge of the complete network would be helpful in identifying the root causes\cite{bexiga2013human,gissen2007cargos}. At a more basic level, having complete pictures of the vesicle traffic system for various organisms and various cell-types would allow comparative studies, and therefore would allow the deduction of modes of evolution of the form of the traffic network, and would also allow us to decipher features such as parts of the traffic system that is unchanging, and therefore likely to be its core, and parts that are more plastic \cite{barlow2017seeing}.

In this paper, using constraints on global vesicle traffic network topology due to local molecular interactions, we take incomplete pictures of vesicle traffic networks as inputs and output various completed versions which can then be tested experimentally.
%

%Previous analyses of VTS have mostly been based on sampling approach~\cite{mani2016wine, mani2016stacking}. 
%%
%In these analyses, vesicle traffic is
%modeled as a dynamical system. 
%%
%The traffic rule specifies how the system transitions from one
%time point to the next. 
%%
%Given a traffic rule and an initial condition, the system is evolved over time until a steady state is reached. By studying a large sample of randomly generated traffic
%rules, such analyses can make statistical claims about vesicle traffic.
%%
%Hence making precise general prediction about the properties of vesicle traffic networks over all possible traffic rules, not just for a sampled subset is not possible.
%%
%SAT and SMT solvers are a perfect fit for this situation, solving combinatorial problems with their exhaustive nature of searching. 
%%
%These tools check a specified logical property
%on every possible state of a model constructed using variables ranging over Boolean or any
%finite discrete type.
